\section{Background}
\label{ch:background}

This chapter is separated into two essential parts which contain background
knowledge necessary to understand the problem and methodology used in this 
thesis. As stated in the introduction the goal of this thesis is to apply deep
learning techniques to the prediction of tweet engagement metrics. Therefore,
the first section (ch. 2.1) presents deep learning basics as well as the current
state-of-the-art of research in this area. In the second section (ch. 2.2), 
basics about social networks are established. The section also covers necessary 
Twitter terminology.

\subsection{Deep Learning}
\label{sec:deep_learning}

The application of deep learning techniques has accelerated the progress of 
machine learning (ML) and artificial intelligence (AI) research in recent years~\cite{Brynjolfsson2017}. 
This section aims to explain deep learning from the ground up. Thus, the 
first subsection (ch. 2.1.1) defines basic terminology around this topic 
and before the second (ch. 2.1.2) delivers historical perspectives on deep learning research. 
Basic elements and algorithms used in deep neural networks are explained in the third 
subsection (ch. 2.1.3), before recent developments in research are
presented (ch. 2.1.4). The last subsection (ch. 2.1.5) describes how
deep learning models are applied in practice.

\input{sec/dl_terminology.tex}

\input{sec/dl_history.tex}

\input{sec/dl_concepts.tex}

\input{sec/dl_developments.tex}

\input{sec/dl_applications.tex}

\subsection{Social Networks}
\label{sec:social_networks}

Applying deep learning to tweet engagement prediction requires a basic
understanding of social network dynamics and structure. Furthermore, a common
understanding of Twitter terminology is necessary for later chapters of this
work. Both areas will be covered in this section, beginning with social
networks in general (ch.~\ref{sub:sn_terminology}) and concluding with
Twitter in detail (ch.~\ref{sub:sn_twitter}).

\input{sec/sn_terminology.tex}

\input{sec/sn_twitter.tex}
